---
title: "A cheetsheet for joining method"
author: "Jason Sun"
email: "sunb@ece.ubc.ca"
date: "`r format(Sys.Date())`"
output: github_document
---

## Intro
This is the cheetsheet for joing method Check this out!~


## Initial Setup
We need to load the library first

```{r}
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(gapminder))
options(geonamesUsername="jasonsunbao")
library(geonames)
library(tidyr)
library(knitr)
library(dplyr)
library(gapminder)
library(reshape2)
```

## Do some work
OK. We will then use the following join methods:

- inner_join(dataset1, dataset2)

- semi_join(dataset1, dataset2)

- left_join(dataset1, dataset2)

- anti_join(dataset1, dataset2)

- inner_join(dataset2, dataset1)

- semi_join(dataset2, dataset1)

- left_join(dataset2, dataset1)

- anti_join(dataset2, dataset1)

- full_join(dataset1, dataset2)

Well, the first dataset will be gapminder and the second dataset will be a manually created dataset with first column as country and second column as capital city.


```{r}
options(geonamesUsername="jasonsunbao")
countryInfo <- GNcountryInfo()
dataset2 <- countryInfo%>%select(countryName, countryCode)
dataset1 <- gapminder
```

Let's try inner_join(), it will get the intersection of the two datasets
```{r}
inner_join(dataset1, dataset2, c("country" = "countryName")) %>% filter(year=="1952")
```

Then Semi_join() will return all rows from x where there are matching values in y, keeping just columns from x
```{r}
semi_join(dataset1, dataset2, c("country" = "countryName")) %>% filter(year=="1952")
```

Then left_join(x, y) will return all rows from x, and all columns from x and y. If there are multiple matches between x and y, all combination of the matches are returned

```{r}
left_join(dataset1, dataset2, c("country" = "countryName")) %>% filter(year=="1952")
```

After that is  anti_join(x, y) will eturn all rows from x where there are not matching values in y, keeping just columns from x, which is a filtering join

```{r}
anti_join(dataset1, dataset2, c("country" = "countryName")) %>% filter(year=="1952")
```

Finally is full_join(x, y) will return all rows and all columns from both x and y. Where there are not matching values, returns NA for the one missing, which is a mutating join

```{r}
full_join(dataset1, dataset2, c("country" = "countryName")) %>% filter(year=="1952")
```






