---
title: "Tidyr_Cheetsheet"
author: "Jason Sun"
email: "sunb@ece.ubc.ca"
date: "`r format(Sys.Date())`"
output: github_document
---

## Intro
This is the cheat sheet for Tidyr. Check this out!~


## Initial Setup
We need to load the library first

```{r}
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(dplyr))
library(tidyr)
library(dplyr)
```

## Do some work
Tidyr is new package that makes it make dataset clean and used to munge with dplyr, ggplot2 and further packages and it has two important properties: 

Each column is a variable.

Each row is an observation.

And it has certain functions for usage:

- gather(): gather() takes multiple columns, and gathers them into key-value pairs: it makes ???wide??? data longer, it has melt(), pivot() and fold()

- spread(): spread() takes two columns (a key-value pair) and spreads them in to multiple columns, making ???long??? data wider.

```{r}
input <- data.frame(
  name = c("Jason", "William", "Tom"),
  a = c(67, 80, 64),
  b = c(56, 90, 50)
)
input
```

Now we want to extend it into a key-value pair. we use gather()
```{r}
input %>% gather(nationality, birthplace, a:b)
```
Hmm. Quite interesting

Let's see how to do that for spread()
```{r}
input %>% gather(nationality, birthplace, a:b) %>% spread(nationality, birthplace)
```

Oh Yeah! It is now returned back to where it looks like for the first place.




