---
title: "A deeper exploration into Gapminder Dataset"
author: "Jason Sun"
email: "sunb@ece.ubc.ca"
date: "`r format(Sys.Date())`"
output: github_document
---



## Intro

Last time during homework 01, I have showed you the most fundmental aspects of gapminder, which is a very basic dataset for entry level people to better explore R. We see the general form of it, which is a two-dimensional table with 6 attributes and 1704 lines of data. Today, we will dig a little deeper as we are becoming more and more professional in R :)

This page will be divided into the following aspects:
- Overall Property of the Dataset
- Variable Property Exploration
- Data Visualization using ggplot2
- Data Filtering with limited conditions to get desired information
- Personal Thought about this work

## Overall Property of Gapminder
We need to load the Gapminder first

```{r}
library(gapminder)
library(tidyverse)
library(ggplot2)
```

So the first question we are going to answer is what is Gapminder? Is it a data.frame, a matrix, a vector, a list?
Well, we can check it by using typeof() function in R. Let's try it out!

```{r}
typeof(gapminder)
```

Wow, it seems it is a list data type.
Then what is its class? 
```{r}
class(gapminder)
```
Hmm, it is a Dataframe.

So next, How many variables/columns do we have? We can use ncol in R
```{r}
ncol(gapminder)
```

Then how many rows/observations do we have? We can use nrow in R
```{r}
nrow(gapminder)
```

So next I will show you some different tricks about how to find out the extent and size of the dataset

To find the length of the dataset
```{r}
length(gapminder)
```

To give a dimension of the dataset
```{r}
dim(gapminder)
```

To give the names of all the columns
```{r}
names(gapminder)
```

Let's have a final statistical overview of the dataset
```{r}
summary(gapminder)
```

Now let's find out the data type of each variable in the dataset
```{r}
typeof(gapminder$country)
typeof(gapminder$continent)
typeof(gapminder$year)
typeof(gapminder$lifeExp)
typeof(gapminder$pop)
typeof(gapminder$gdpPercap)
```

Next we will explore one categorical variable and one quantative variable.
The categorical data we pick up will be the year variable
Let's get a summary of year first:
```{r}
summary(gapminder$year)
table(gapminder$year)
```
So it is quite clear that the year ranges from 1952 to 2007 and median value is 1980 and mean value is also 1980

Let's also take a look at the quantative variable lifeExp
```{r}
summary(gapminder$lifeExp)
hist(gapminder$lifeExp)
```
From the histgram we can see that the 70~75 takes up most of the life expectency and max lifeExp is 82.6 and Min is 23.6 where median is 60.71.

## Plot Exploration with ggplot2 for Gapminder 
Let's try to see the scatterplot for the relationship between lifeExp and gdpPercap
```{r}
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp)) + geom_point()
```

Let's try to find out the log10(gdpPerCap) towards lifeExp
```{r}
ggplot(gapminder, aes(x = log10(gdpPercap), y = lifeExp)) + geom_point()
```

Let's try to plot a quantative variable lifeExp and make a histogram of it
```{r}
hist(gapminder$lifeExp)
```

Now Let's try something interesting. How about we draw a plot by year with different countries and their respective lifeExp? Would that be difficult? 
```{r}
jCountries <- c("Canada", "Rwanda", "Cambodia", "Mexico")
ggplot(subset(gapminder, country %in% jCountries), aes(x = year, y = lifeExp, color = country)) + geom_line() + geom_point()
```

## Filtering and Selecting
Sometimes we don't need to use all the data provided. So we will need to filter and select the data that we want. R has provided us with some really useful functions that can easily help to do the filtering. Let's take some real examples.
For example, if we want to plot a lifeExp graph with ony country of Canada in ggplot2, how could we do that? We can add a parameter called filter in ggplot to assgin the country to Canada then that's it!
```{r}
suppressPackageStartupMessages(library(dplyr))
ggplot(gapminder %>% filter(country == "Canada"), aes(x = year, y = lifeExp)) + geom_line() + geom_point()
```

Cool! So how about using the filter() and select() together to plot the lifeExp by year for Rwanda? Let's try this:
```{r}
filter(gapminder, country == c("Rwanda")) %>%select(year, lifeExp) %>%ggplot(aes(x=year, y=lifeExp))+ geom_line() + geom_point()
```

Wops, there seems to be a drastic drop in 1991 where the lifeExp is below 25, which I think may be a mistake I guess..

## Some thoughts
Homework2 is a little harder than homework1 with more things to consider and requires you more proficient in R for data types, plots, common operations, etc. Afterall, it is not so difficult as thought and we can make things better in the next homework assignment.





















