---
title: "omdb_dataset_download_show"
author: "Jason Sun"
date: "12/5/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Let's load some library first

```{r cars}
library(purrr)
library(ggplot2)
library(reshape)
library(stringr)
library(tidyverse)
library(glue)
library(plyr)
```
## Let's see the requirement for this work
Create a dataset with multiple records by requesting data from an API using the httr package.

GET() data from the API and convert it into a clean and tidy data frame. Store that as a file ready for (hypothetical!) downstream analysis. Do just enough basic exploration of the resulting data, possibly including some plots, that you and a reader are convinced you’ve successfully downloaded and cleaned it.

Take as many of these opportunities as you can justify to make your task more interesting and realistic, to use techniques from elsewhere in the course (esp. nested list processing with purrr), and to gain experience with more sophisticated usage of httr.

Get multiple items via the API (i.e. an endpoint that returns multiple items at once) vs. use an iterative framework in R.
Traverse pages.

Send an authorization token. The GitHub API is definitely good one to practice with here.

Use httr’s facilities to modify the URL and your request, e.g., query parameters, path modification, custom headers.

## Work That I have Done
My work contains three part: Fetch Data, Plot and Save and finally show the result here.

I used the rquired httr package to download the data via OMDBApi and automate the whole process using Makefile.

I created a function called get_movie_rating(name), which can get the rating of a specific movie with input name and the rating contains IMDB, RottonTomato and Met.

After calling the function, data will be saved as CSV file to local disk and during the plot process, we will do various filtering and take out the top 5 score of IMDB, RT and Met for Star Trek and Transformer as example and show as table as well as bar chart for the result to give the user an overview of what is the best movie in the series of Transformer and Star Trek.

The whole process is like the following: 

Fectch Data with OMDBApi and httr ----> Parse, Filter and Plot Graph ----> Show The Result

## Let's do some work

You can also embed plots, for example:

```{r,warning = FALSE, message=FALSE}
#Get the data from the csv
movie_rating_transformers <-read_csv("/Users/jasonsun/stat545/work1/STAT545-hw01--Bao-Sun/hw10/OMDB/transformers_rating.csv")
movie_rating_startrek <- read_csv("/Users/jasonsun/stat545/work1/STAT545-hw01--Bao-Sun/hw10/OMDB/star_trek_rating.csv")
```


Let's get a brief overview of the current dataset we have for transformers
```{r,warning = FALSE, message=FALSE}
knitr::kable(movie_rating_transformers)
```

Let's get a brief overview of the current dataset we have for star trek
```{r,warning = FALSE, message=FALSE}
knitr::kable(movie_rating_startrek)
```

Let's see top 5 for IMDB for transformers
```{r,warning = FALSE, message=FALSE}
movie_rating_transformers_temp <- movie_rating_transformers  %>% select(movie_title, imdb)%>% filter(imdb > 0) %>% arrange(desc(imdb))
movie_rating_transformers_temp <- movie_rating_transformers_temp[0:5,]
knitr::kable(movie_rating_transformers_temp)
```

Let's see top 5 for RT Score for transformers
```{r,warning = FALSE, message=FALSE}
movie_rating_transformers_temp_rt <- movie_rating_transformers  %>% select(movie_title, RT)%>% filter(RT > 0) %>% arrange(desc(RT))#Let's choose top 5 for RT for transformers
movie_rating_transformers_temp_rt <- movie_rating_transformers_temp_rt[0:5,]
knitr::kable(movie_rating_transformers_temp_rt)
```

Let's see top 5 for Met Score for transformers
```{r,warning = FALSE, message=FALSE}
movie_rating_transformers_temp_met <- movie_rating_transformers  %>% select(movie_title, Met)%>% filter(Met > 0) %>% arrange(desc(Met))#Let's choose top 5 for RT for transformers
movie_rating_transformers_temp_met <- movie_rating_transformers_temp_met[0:5,]
knitr::kable(movie_rating_transformers_temp_met)
```

Let's see top 5 for IMDB Scorefor Star Trek
```{r,warning = FALSE, message=FALSE}
movie_rating_star_trek_temp <- movie_rating_startrek  %>% select(movie_title, imdb)%>% filter(imdb > 0) %>% arrange(desc(imdb))
#Let's choose top 5 for IMDB for Star Trek
movie_rating_star_trek_temp <- movie_rating_star_trek_temp[0:5,]
knitr::kable(movie_rating_star_trek_temp)
```

Let's see top 5 for RT Score for Star Trek
```{r,warning = FALSE, message=FALSE}
movie_rating_star_trek_temp_rt <- movie_rating_startrek  %>% select(movie_title, RT)%>% filter(RT > 0) %>% arrange(desc(RT))
movie_rating_star_trek_temp_rt <- movie_rating_star_trek_temp_rt[0:5,]
#Then let's have an overview
knitr::kable(movie_rating_star_trek_temp_rt)
```

Let's see top 5 for Met Score for Star Trek
```{r,warning = FALSE, message=FALSE}
movie_rating_star_trek_temp_met <- movie_rating_startrek  %>% select(movie_title, Met)%>% filter(Met > 0) %>% arrange(desc(Met))
#Let's choose top 5 for Met
movie_rating_star_trek_temp_met <- movie_rating_star_trek_temp_met[0:5,]
#Then let's have an overview
knitr::kable(movie_rating_star_trek_temp_met)
```

Let's draw the graph for all of them

- Let's see top 5 for IMDB for transformers

- Let's see top 5 for RT Score for transformers

- Let's see top 5 for Met Score for transformers

- Let's see top 5 for IMDB Scorefor Star Trek

- Let's see top 5 for RT Score for Star Trek

- Let's see top 5 for Met Score for Star Trek




